---
title: "WUP_04"
author: "Jonathan Urbano"
date: "June 12, 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
require(stats)
thedata<-read.csv(file="GSScsv.csv")
```

```{r eval=FALSE}
thedata
require(tidyverse)
require(tigerstats)
thedata<-read.csv(file="GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

We wanted to find out by looking at the data what sort of data we had. We copied write up 3 to read the data in and changed the character variables into and factor variables. We saw that we had a numerical data which was WORDSUM and we now have a factor variable called DEGREE. SO we deduced that we have a numerical `twiddle` categorical problem. 

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
ggplot(thedata, aes(x = DEGREE, y = WORDSUM)) +
geom_boxplot(fill = "skyblue", color = "darkblue", outlier.shape = NA) +
theme_minimal() +
labs(x = "DEGREE", y = "WORDSUM") +
theme(plot.title = element_text(hjust = 0.5)) +
ggtitle("Fancy Box Plot")
```

LT High School would not overlap with Junior College or Graduate, thus knowing that it is very likely to be different schooling. In addition, graduate does not overlap with High School. Therefore, it is also likely to be different. Ultimately, LT High School and Graduate are not only different from one another they are both different that another type of schooling, such as High School and Junior College.

### Numerical Descriptive Results

```{r}
favstats(WORDSUM~DEGREE,data=thedata)
```
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

It is evident that the graduates have the highest median among all other categories. In addition, the third quartile values of high school, students and light high school students are relatively close and smaller than all the other values. The light high school students have the overall lowest median and mean values 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
# Perform one-way ANOVA
model <- aov(WORDSUM ~ DEGREE, data = thedata)

# Conduct post-hoc test (Tukey's HSD)
posthoc <- TukeyHSD(model)

# Print the results
print(posthoc)
```

```{r}
# Assuming you have the "thedata" data frame with "DEGREE" and "WORDSUM" columns

# Load the emmeans package
library(emmeans)

# Perform one-way ANOVA
result <- aov(WORDSUM ~ DEGREE, data = thedata)

# Perform Scheffe's post hoc test
comparison <- emmeans(result, pairwise ~ DEGREE, adjust = "scheffe")

# Print the comparison table
print(comparison)
```

The high school-graduate and the light high school graduate lines didn't cross the zero line, meaning they don't have zero in their confidence difference so there's no difference between these two difference categories. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
